Decompositions of congruence subgroups of Chevalley groups

### 2015-10-11 congrsubgr-boundgen.tex
Commas spaces in Gauss, DV, CB and SF

\textsc{Case 1:} $\alpha\geqslant\beta$. 
Since $p^{\alpha-\beta}a\perp bm^2$ and $p\perp bm^2$, there exist infinitely many rational primes $q$ of the form $p^{\alpha-\beta}a+bm^2k$,
such that $p$ is a primitive root modulo $q$. 
Without loss of generality we may assume that $q$ is prime to $b$. 
Write
\begin{align*}
g_1 & = x_{12}(mk)\cdot g =
\begin{pmatrix} p^\beta q & * \\ p^\beta bm & * \end{pmatrix}. \\
\intertext{
There exists $u\geqslant 1$ such that $p^u\equiv b\pmod q$, say $p^u=b+lq$. Then
}
g_2 & = x_{21}(ml)\cdot g_1 =
\begin{pmatrix} p^\beta q & * \\ mp^{\beta+u} & * \end{pmatrix}. \\
\intertext{
Since $g_2\equiv 1\pmod m$, we can write $p^\beta q=1+cm$ for some $c$. Now let
}
g_3 & = x_{12}\left(\dfrac{-c}{p^{\beta+u}}\right)\cdot g_2 =
\begin{pmatrix} 1 & * \\ mp^{\beta+u} & * \end{pmatrix}, \\
g_4 & = x_{21}\left(-mp^{\beta+u}\right)\cdot g_3 =
\begin{pmatrix} 1 & * \\ 0 & * \end{pmatrix}, \\
g_5 & = x_{12}\left(\dfrac{c}{p^{\beta+u}}\right)\cdot g_4 =
\begin{pmatrix} 1 & * \\ 0 & * \end{pmatrix}.
\end{align*}
Notice that $g_5=z_{21}\left(-mp^{\beta+u},c/p^{\beta+u}\right)\cdot g_2$ hence $g=x_{12} \cdot x_{21} \cdot z_{21} \cdot x_{12}$
and the length of $g$ does not exceed $4$.

@@

%Now it is easy to derive bounded generation property for $\G(\Phi,\mathcal{O}_S,I)$ from the above statements.
\begin{lemma}\label{lemma:width-dedekind}
Let $\Phi$ be an irreducible classical root system of rank $\ell \geqslant 2$.
If $k$ has a real embedding, then $\G(\Phi,\mathcal{O}_S,I)$ has finite width with respect to the generating set $\mathcal{Z}(\Sigma_\ell)$.
\end{lemma}
\begin{proof}
First of all, notice that $\asr(I) \leq \asr(\mathcal{O}_S) \leq 2$. 
By \cref{cor:bass-kolster-iterated} we can present any element 
of $G=\G(\Phi, \mathcal{O}_S, I)$ as a product of a finite number of generators $\mathcal{Z}(\Sigma_\ell)$ and one element of 
$\G(\{\alpha_\ell, -\alpha_\ell\}, \mathcal{O}_S, I)\cong\SL(2,\mathcal{O}_S,I)$.
Consequently, to prove the statement of the lemma it suffices to express every element 
$g = \begin{psmallmatrix}1+a & b \\ c & 1+d \end{psmallmatrix} \in G_0$
as a product of a finite number of generators contained in some rank $2$ subgroup of $G$ containing $G_0$.

From $\det(g)=1$ we conclude that $a+d=bc-ad\in I^2$. 
Recall that Vaserstein's congruence subgroup is defined as
\[ G(I,I)=\left\{ \begin{pmatrix}1+a & b \\ c & 1+d\end{pmatrix}\in\SL(2,\mathcal{O}_S)\;\middle|\; a,d\in I^2,\ b,c\in I \right\}. \]
Notice that $g_1=g\cdot z_{21}(a,1)$ is contained in $G(I,I)$, indeed
\[ \begin{pmatrix} 1+a & b \\ c & 1+d \end{pmatrix} \cdot \begin{pmatrix} 1-a & -a \\ a & 1+a \end{pmatrix} = \begin{pmatrix} 1+ba-a^2 & b-a-ba-a^2 \\ c+a+ad-ac & 1+bc-ac \end{pmatrix} \in G(I,I). \]
For any matrix $g'=\begin{psmallmatrix}1+a & b \\ c & 1+d\end{psmallmatrix}\in G(I,I)$ the matrix $x_{21}(-c)\cdot g'\cdot x_{12}(-b)$ lies in $\SL\left(2,\mathcal{O}_S,I^2\right)$.

By \cref{lemma:Stepanov-ideal} the group $\E\left(\Phi,\mathcal{O}_S,I^2\right)$ is contained in $\E(\Phi,I)$ for any root system $\Phi\neq\rC_\ell$ of rank $\geqslant2$.
Notice that under the assumptions of the lemma it is known that $\E(\Phi,I)$ has finite width with respect to $\mathcal{X}$, see~\cite{Tavgen93} or \cite[Theorem~3.3]{TavgenThesis}.

In remains to consider the case $\Phi=\rC_\ell$. First of all, notice that $2abc-abd\in II^{\indexbox{2}}$, indeed
\[ \det(g_1)=a^3d-3a^2bc+a^2bd+ab^2c+a^3+a^2b+a^2d+2abc-abd+1. \]
Consequently, we have that
\[ g_2=x_{21}(-a-c)\cdot g_1\cdot x_{12}(a-b)\equiv
\begin{pmatrix}
1+ab-a^2 & -ab-a^2 \\ ad-ac-abc & 1-ab+a^2
\end{pmatrix}\bmod II^{\indexbox{2}}. \]
Now for $g_3=g_2\cdot z_{12}\left(a^2-ab,1\right)$ we have that
\begin{align*}
& g_3\equiv\begin{pmatrix} 1 & -2ab \\ -abc-a^2+ab-ac+ad & 1 \end{pmatrix}\bmod II^{\indexbox{2}},\\
& g_4=x_{12}(2ab)\cdot g_3\equiv x_{21}\left(-abc-a^2+ab-ac+ad\right)\bmod II^{\indexbox{2}}.
\end{align*}
Thus $g_4\cdot x_{21}(*)\in\SL\left(2,\mathcal{O}_S,II^{\indexbox{2}}\right)$ is contained in $\E(\rC_\ell,I)$ by \cref{lemma:Stepanov-ideal} and therefore can be expressed as a bounded product of $x_\alpha$.
\end{proof}

\begin{proof}[Proof of \cref{thm:width}] The second assertion of the theorem is contained in \cref{lemma:srRI1_width}.
The the first and the third assertions follow from \cref{lemma:Z1p,lemma:width-dedekind}. \end{proof}

>>>>

\textsc{Case 1:} $\alpha\geqslant\beta$. 
Since $p^{\alpha-\beta}a\perp bm^2$ and $p\perp bm^2$, there exist infinitely many rational primes $q$ of the form $p^{\alpha-\beta}a+bm^2k$,
such that $p$ is a primitive root modulo $q$. 
Without loss of generality we may assume that $q$ is prime to $b$. 
Write
\begin{align*}
g_1 & = x_{12}(mk)\cdot g =
\begin{pmatrix} p^\beta q & * \\ p^\beta bm & * \end{pmatrix}. \\
\intertext{
There exists $u\geqslant 1$ such that $p^u\equiv b\pmod q$, say $p^u=b+lq$. Then
}
g_2 & = x_{21}(ml)\cdot g_1 =
\begin{pmatrix} p^\beta q & * \\ mp^{\beta+u} & * \end{pmatrix}. \\
\intertext{
Since $g_2\equiv 1\pmod m$, we can write $p^\beta q=1+cm$ for some $c$. Now let
}
g_3 & = x_{12}\left(\dfrac{-c}{p^{\beta+u}}\right)\cdot g_2 =
\begin{pmatrix} 1 & * \\ mp^{\beta+u} & * \end{pmatrix}, \\
g_4 & = x_{21}\left(-mp^{\beta+u}\right)\cdot g_3 =
\begin{pmatrix} 1 & * \\ 0 & * \end{pmatrix}, \\
g_5 & = x_{12}\left(\dfrac{c}{p^{\beta+u}}\right)\cdot g_4 =
\begin{pmatrix} 1 & * \\ 0 & * \end{pmatrix}.
\end{align*}
Notice that $g_5=z_{21}\left(-mp^{\beta+u}, c/p^{\beta+u}\right)\cdot g_2$ hence $g=x_{12} \cdot x_{21} \cdot z_{21} \cdot x_{12}$
and the length of $g$ does not exceed $4$.

@@

%Now it is easy to derive bounded generation property for $\G(\Phi, \mathcal{O}_S, I)$ from the above statements.
\begin{lemma}\label{lemma:width-dedekind}
Let $\Phi$ be an irreducible classical root system of rank $\ell \geqslant 2$.
If $k$ has a real embedding, then $\G(\Phi, \mathcal{O}_S, I)$ has finite width with respect to the generating set $\mathcal{Z}(\Sigma_\ell)$.
\end{lemma}
\begin{proof}
First of all, notice that $\asr(I) \leq \asr(\mathcal{O}_S) \leq 2$. 
By \cref{cor:bass-kolster-iterated} we can present any element 
of $G=\G(\Phi, \mathcal{O}_S, I)$ as a product of a finite number of generators $\mathcal{Z}(\Sigma_\ell)$ and one element of 
$\G(\{\alpha_\ell, -\alpha_\ell\}, \mathcal{O}_S, I)\cong\SL(2, \mathcal{O}_S, I)$.
Consequently, to prove the statement of the lemma it suffices to express every element 
$g = \begin{psmallmatrix}1+a & b \\ c & 1+d \end{psmallmatrix} \in G_0$
as a product of a finite number of generators contained in some rank $2$ subgroup of $G$ containing $G_0$.

From $\det(g)=1$ we conclude that $a+d=bc-ad\in I^2$. 
Recall that Vaserstein's congruence subgroup is defined as
\[ G(I, I)=\left\{ \begin{pmatrix}1+a & b \\ c & 1+d\end{pmatrix}\in\SL(2, \mathcal{O}_S)\;\middle|\; a, d\in I^2, \ b, c\in I \right\}. \]
Notice that $g_1=g\cdot z_{21}(a, 1)$ is contained in $G(I, I)$, indeed
\[ \begin{pmatrix} 1+a & b \\ c & 1+d \end{pmatrix} \cdot \begin{pmatrix} 1-a & -a \\ a & 1+a \end{pmatrix} = \begin{pmatrix} 1+ba-a^2 & b-a-ba-a^2 \\ c+a+ad-ac & 1+bc-ac \end{pmatrix} \in G(I, I). \]
For any matrix $g'=\begin{psmallmatrix}1+a & b \\ c & 1+d\end{psmallmatrix}\in G(I, I)$ the matrix $x_{21}(-c)\cdot g'\cdot x_{12}(-b)$ lies in $\SL\left(2, \mathcal{O}_S, I^2\right)$.

By \cref{lemma:Stepanov-ideal} the group $\E\left(\Phi, \mathcal{O}_S, I^2\right)$ is contained in $\E(\Phi, I)$ for any root system $\Phi\neq\rC_\ell$ of rank $\geqslant2$.
Notice that under the assumptions of the lemma it is known that $\E(\Phi, I)$ has finite width with respect to $\mathcal{X}$, see~\cite{Tavgen93} or \cite[Theorem~3.3]{TavgenThesis}.

In remains to consider the case $\Phi=\rC_\ell$. First of all, notice that $2abc-abd\in II^{\indexbox{2}}$, indeed
\[ \det(g_1)=a^3d-3a^2bc+a^2bd+ab^2c+a^3+a^2b+a^2d+2abc-abd+1. \]
Consequently, we have that
\[ g_2=x_{21}(-a-c)\cdot g_1\cdot x_{12}(a-b)\equiv
\begin{pmatrix}
1+ab-a^2 & -ab-a^2 \\ ad-ac-abc & 1-ab+a^2
\end{pmatrix}\bmod II^{\indexbox{2}}. \]
Now for $g_3=g_2\cdot z_{12}\left(a^2-ab, 1\right)$ we have that
\begin{align*}
& g_3\equiv\begin{pmatrix} 1 & -2ab \\ -abc-a^2+ab-ac+ad & 1 \end{pmatrix}\bmod II^{\indexbox{2}}, \\
& g_4=x_{12}(2ab)\cdot g_3\equiv x_{21}\left(-abc-a^2+ab-ac+ad\right)\bmod II^{\indexbox{2}}.
\end{align*}
Thus $g_4\cdot x_{21}(*)\in\SL\left(2, \mathcal{O}_S, II^{\indexbox{2}}\right)$ is contained in $\E(\rC_\ell, I)$ by \cref{lemma:Stepanov-ideal} and therefore can be expressed as a bounded product of $x_\alpha$.
\end{proof}

\begin{proof}[Proof of \cref{thm:width}] The second assertion of the theorem is contained in \cref{lemma:srRI1_width}.
The the first and the third assertions follow from \cref{lemma:Z1p, lemma:width-dedekind}. \end{proof}

### 2015-10-11 dennis-vaserstein.tex
Commas spaces in Gauss, DV, CB and SF

%The relative elementary subgroup $\E(\Phi, R, I)$ admits Dennis---Vaserstein decomposition, i.\,e. $\E(\Phi, R, I) = A_{rs}$ under the following assumptions on $I$ and $\Phi$.
%\[\begin{array}{l@{\qquad}c@{\quad}c@{\qquad}l}
%\Phi                                 & s    & r      & \text{ring condition} \\ \hline\vphantom{\Bigl(}
%\rA_\ell,\ \ell\geqslant 2           & 1    & \ell   & \sr(I) \leqslant \ell-1 \\    
%\rB_\ell,\ \ell\geqslant 2           & 1    & \ell   & \sr(I) \leqslant \ell-1 \\
%\rC_\ell,\ \ell \geqslant 2          & 1    & \ell   & \sr(I) \leqslant \ell-1 \\
%\rD_\ell,\ \ell\geqslant 4           & 1    & \ell   & \sr(I) \leqslant \ell-2 \\ 
%\rD_\ell,\ \ell\geqslant 4           & \ell & \ell-1 & \sr(I) \leqslant 2      \\
%\rE_\ell,\ \ell=6,7                  & \ell & 2      & \sr(I) \leqslant \ell-3 \\ 
%\rE_\ell,\ \ell=6,7                  & \ell & 1      & \asr(I)\leqslant \ell-2 \\ 
%\rF_4                                & 4    & 1      & \asr(I)\leqslant 3 \end{array}\]
Denote by $N_{rs}$ the submonoid of $\E(\Phi, R)$ consisting of elements which normalize the subset $A_{rs}$, i.\,e.
\[ N_{rs} = \left\{ g\in \E(\Phi, R)\ \middle|\ g \cdot A_{rs} \cdot g^{-1} \subseteq A_{rs} \right\}. \]

\begin{lemma}\label{lemma:dv-normal} 
For every $\alpha \in \Delta_{\{r,s\}} \cup (\Phi^+ \setminus (\Sigma^+_r \cap \Sigma^+_s))$ one has $X_\alpha(R) \subseteq N_{rs}$. \end{lemma}
\begin{proof}
Notice that for every $i$ the group $\EP_i(R, I)$ is normalized by $\EP_i(R)$, hence a fortiori it is normalized by $X_\alpha(R)$, $\alpha \in S_i^+$.
Since $\U(\Sigma_r^- \cap \Sigma_s^-, I)$ is normalized by $\E(\Delta_{\{r,s\}}, R)$ we obtain the assertion of the lemma for $\alpha \in \Delta_{\{r, s\}}$.

@@

\begin{proof}[Proof of \cref{thm:DennisVaserstein}]
First of all, notice that in the case $\Phi=\rG_2$ the assertion of the theorem formally follows from \cref{thm:srRI1} and identities~\ref{rel:h-comm}--\ref{rel:h-w}.
Since $A_{rs} = \E(\Phi, R, I)$ implies $A_{sr} = \E(\Phi, R, I)$, it suffices to consider only the following possibilities for $\Phi$, $s$, $r$.
\begin{table}[htb]
\[\begin{array}{l @{\qquad} l @{\qquad} c @{\quad} c @{\quad} c @{\qquad} c @{\qquad} c @{\qquad} c}
\Phi                                 & s    &r      & |\Lambda(\pi)| & \text{type of $\pi$} & \text{type of $\Delta_r$} & |\Lambda(\pi')|& |\Gamma|  \\ \hline\vphantom{\Bigl(}
\rA_\ell,\ \ell\geqslant 2           & 1    &\ell   & \ell+1         & \text{natural}       & \rA_{\ell-1}              & \ell           & 1  \\     
\rB_\ell,\ \ell\geqslant 2           & 1    &\ell   & 2\ell+1        & \text{natural}       & \rA_{\ell-1}              & \ell           & 1  \\     
\rC_\ell,\ \ell\geqslant 2           & 1    &\ell   & 2\ell          & \text{natural}       & \rA_{\ell-1}              & \ell           & 1  \\
\rD_\ell,\ \ell\geqslant 4           & 1    &\ell   & 2\ell          & \text{natural}       & \rA_{\ell-1}              & \ell           & 2  \\ 
\rD_\ell,\ \ell\geqslant 4           & \ell &\ell-1 & 2^{\ell-1}     & \text{half-spinor}   & \rA_{\ell-1}              & \ell           & \ell-2  \\
\rE_\ell,\ \ell=6,7                  & \ell &2      & 27, 56         & \text{minimal}       & \rA_{\ell-1}              & \ell           & 3       \\ 
\rE_\ell,\ \ell=6,7                  & \ell &1      & 27, 56         & \text{minimal}       & \rD_{\ell-1}              & 2(\ell-1)      & \ell-1  \\
\rF_4                                & 4    &1     & 26             & \text{minimal}       & \rC_3                     & 6              & 3\end{array}\]
 \caption{List of the cases considered in the proof of \cref{thm:DennisVaserstein}.} \label{table:dv-reps}
\end{table}

Denote by $L_{rs}$ the set consisting of all the elements $g\in \E(\Phi, R, I)$ such that $g \cdot A_{rs} \subseteq A_{rs}$.
It is easy to see that $L_{rs}$ contains $\EP_r(R, I)$ and is normalized by $N_{rs}$, i.\,e. ${L_{rs}}^{N\ssub{rs}}\subseteq L_{rs}$. Indeed, for $g\in N_{rs}$, $h\in L_{rs}$ one has
\begin{equation}\label{rel:NnormL} h^g \cdot A_{rs} = g^{-1} \cdot h \cdot g \cdot A_{rs} \subseteq g^{-1} \cdot h \cdot A_{rs} \cdot g \subseteq {A_{rs}}^g \subseteq A_{rs}.\end{equation}

@@

The next goal is to demonstrate the inclusion $\U(\Sigma_s^+, R) \subseteq N_{rs}$ (which implies $\EP_s(R) \subseteq N_{rs}$).
All the possibilities for $\Phi$, $s$ and $r$ fall into one of the following three cases:
\begin{enumerate}
 \item \textit{Case $m_s(\alpha_{\mathrm{max}})=1$ and $\Phi$ is simply laced.}
  In this case all the roots of $\Sigma^+_s$ have the same $s$-shape and length, hence the assertion follows from \cref{item-trans2}, \eqref{rel:NnormL} and the fact that we already have the inclusion $X_{\alpha_s}(R) \subseteq N_{rs}$.
 \item \textit{Case $\Phi = \rB_\ell$.} Using the argument from the previous case we get that $N_{rs}$ contains the subgroup generated by the root subgroups $X_\alpha(R)$, where $\alpha$ varies over long roots of $\Sigma_s^+$. 
  It remains to prove the inclusion $X_{1,0}(R) \subset N_{rs}$.
  Specializing identity~\eqref{rel:CCF} we get \begin{equation*} \label{rel:CCF-specBC} x_{1,0}(ab) = [x_{1, 2}(a), x_{2, 0}(b)] \cdot x_{1,-2}(-a b^2) \end{equation*}
  Since $X_{2, 0}(R)$ and $X_{1,-2}(R)$ are contained in $N_{rs}$, the inclusion $X_{1,0}(R) \subseteq N_{rs}$ follows.
 \item \textit{Case $\Phi = \rC_\ell, \rF_4$.}
 We settle these remaining cases in \cref{lemma:DVST} where we invoke the stability conditions one more time. 
 In fact, we prove a stronger result, namely that $N_{rs}$ contains $X_{-\alpha\ssub{s}}(R)$ and, consequently, $N_{rs} = \E(\Phi, R)$.
\end{enumerate}

@@

The last inclusion follows from the same transitivity argument as above and the fact that $X_\alpha(I) \subseteq \EP_r(R, I) \subseteq L_{rs}$ for $\alpha \in \Delta_r \cup \Sigma_r$.
For $\Phi=\rB_\ell$ one has to once again apply Chevalley commutator formula~\eqref{rel:CCF}:
\[ X_{0,1}(I) \subseteq X_{2,1}(I) \cdot X_{2,1}(I)^{x_{0,2}(1)} \cdot X_{-2,1}(I) \subseteq L_{rs} \cdot L_{rs}^{N_{rs}} \cdot L_{rs} \subseteq L_{rs}. \qedhere \] 
\end{proof}

@@

\begin{lemma}\label{lemma:DVST}
Assume that one of the following holds:
\begin{lemlist}
 \item \label{lemma:DVcaseCl} $\Phi=\rC_\ell$ and $\sr(I) \leq \ell-1$;
 \item \label{lemma:DVcaseF4} $\Phi=\rF_4$ and $\asr(I) \leq 3$.
\end{lemlist}
Then one has $X_{-\alpha_s}(R) \subseteq N_{rs}$.
\end{lemma}
\begin{proof}
Consider the subset $C \subseteq \EP_s(R, I)$ consisting of elements satisfying the following condition:
\begin{itemize}
 \item \textsc{Case $\Phi=\rC_\ell$.} Matrix entries $(g_{i,2})$, $i=2,\ldots, \ell$ form an $I$-unimodular column of height $\ell-1$.
 \item \textsc{Case $\Phi=\rF_4$.} Matrix entries $(g_{\lambda, \lambda}, \ldots, g_{\lambda, \lambda - \alpha_3 - \alpha_2 - \alpha_1})$, where $\lambda = \varpi_4 - \alpha_4$ (see~\cite[Fig.~26]{PSV98}),
                                     form an $I$-unimodular column of height $4$. \end{itemize}

@@

By the very definition of $C$, for every $g \in C$ one can choose $y \in \U(\Sigma_s^+ \cap \Delta_r, I)$ such that $(y \cdot g)_{\varpi_s,\varpi_s - \alpha_s} = 0$.
Consequently, for every $g\in C$ one has
\begin{multline*}
 \EP_r(R, I) \cdot \U(\Sigma_s^- \cap \Sigma_r^-, I) \cdot g \subseteq \EP_r(R, I) y^{-1} \cdot \U(\Sigma_r^-\cap \Sigma_s^-, I)^{y^{-1}} \cdot y g \subseteq \\
  \subseteq \EP_r(R, I) \cdot \U(\Sigma_r^-, I) \cdot y g
\end{multline*}
Notice that the matrix entry $(yg)_{\varpi_s,\varpi_s}$ is invertible.
From the choice of $y$ it follows that for every $\xi\in R$ the element $g_1:=(yg)^{x_{-\alpha_s}(\xi)}$
satisfies the assumptions of \cref{lemma:Chevalley-Matsumoto} and therefore can be rewritten as $g_1 = uh$ for some $u \in \U(\Sigma_s^-, I)$, $h \in \EP_1(R, I)$.
Consequently, one has
\begin{multline*} {A_{rs}}^{X_{-\alpha_{s}}(R)} \subseteq \EP_r(R, I)^{X_{-\alpha_{s}}(R)} \cdot \U(\Sigma_r^-, I)^{X_{-\alpha_{s}}(R)} \cdot \U(\Sigma_s^-, I) \cdot \EP_s(R, I) \subseteq \\
 \subseteq \EP_\ell(R, I) \cdot \U(\Phi^-, I) \cdot \EP_s(R, I) \subseteq A_{rs} \end{multline*}
 as claimed.
\end{proof}

>>>>

%The relative elementary subgroup $\E(\Phi, R, I)$ admits Dennis---Vaserstein decomposition, i.\, e. $\E(\Phi, R, I) = A_{rs}$ under the following assumptions on $I$ and $\Phi$.
%\[\begin{array}{l@{\qquad}c@{\quad}c@{\qquad}l}
%\Phi                                 & s    & r      & \text{ring condition} \\ \hline\vphantom{\Bigl(}
%\rA_\ell, \ \ell\geqslant 2           & 1    & \ell   & \sr(I) \leqslant \ell-1 \\    
%\rB_\ell, \ \ell\geqslant 2           & 1    & \ell   & \sr(I) \leqslant \ell-1 \\
%\rC_\ell, \ \ell \geqslant 2          & 1    & \ell   & \sr(I) \leqslant \ell-1 \\
%\rD_\ell, \ \ell\geqslant 4           & 1    & \ell   & \sr(I) \leqslant \ell-2 \\ 
%\rD_\ell, \ \ell\geqslant 4           & \ell & \ell-1 & \sr(I) \leqslant 2      \\
%\rE_\ell, \ \ell=6, 7                  & \ell & 2      & \sr(I) \leqslant \ell-3 \\ 
%\rE_\ell, \ \ell=6, 7                  & \ell & 1      & \asr(I)\leqslant \ell-2 \\ 
%\rF_4                                & 4    & 1      & \asr(I)\leqslant 3 \end{array}\]
Denote by $N_{rs}$ the submonoid of $\E(\Phi, R)$ consisting of elements which normalize the subset $A_{rs}$, i.\, e.
\[ N_{rs} = \left\{ g\in \E(\Phi, R)\ \middle|\ g \cdot A_{rs} \cdot g^{-1} \subseteq A_{rs} \right\}. \]

\begin{lemma}\label{lemma:dv-normal} 
For every $\alpha \in \Delta_{\{r, s\}} \cup (\Phi^+ \setminus (\Sigma^+_r \cap \Sigma^+_s))$ one has $X_\alpha(R) \subseteq N_{rs}$. \end{lemma}
\begin{proof}
Notice that for every $i$ the group $\EP_i(R, I)$ is normalized by $\EP_i(R)$, hence a fortiori it is normalized by $X_\alpha(R)$, $\alpha \in S_i^+$.
Since $\U(\Sigma_r^- \cap \Sigma_s^-, I)$ is normalized by $\E(\Delta_{\{r, s\}}, R)$ we obtain the assertion of the lemma for $\alpha \in \Delta_{\{r, s\}}$.

@@

\begin{proof}[Proof of \cref{thm:DennisVaserstein}]
First of all, notice that in the case $\Phi=\rG_2$ the assertion of the theorem formally follows from \cref{thm:srRI1} and identities~\ref{rel:h-comm}--\ref{rel:h-w}.
Since $A_{rs} = \E(\Phi, R, I)$ implies $A_{sr} = \E(\Phi, R, I)$, it suffices to consider only the following possibilities for $\Phi$, $s$, $r$.
\begin{table}[htb]
\[\begin{array}{l @{\qquad} l @{\qquad} c @{\quad} c @{\quad} c @{\qquad} c @{\qquad} c @{\qquad} c}
\Phi                                 & s    &r      & |\Lambda(\pi)| & \text{type of $\pi$} & \text{type of $\Delta_r$} & |\Lambda(\pi')|& |\Gamma|  \\ \hline\vphantom{\Bigl(}
\rA_\ell, \ \ell\geqslant 2           & 1    &\ell   & \ell+1         & \text{natural}       & \rA_{\ell-1}              & \ell           & 1  \\     
\rB_\ell, \ \ell\geqslant 2           & 1    &\ell   & 2\ell+1        & \text{natural}       & \rA_{\ell-1}              & \ell           & 1  \\     
\rC_\ell, \ \ell\geqslant 2           & 1    &\ell   & 2\ell          & \text{natural}       & \rA_{\ell-1}              & \ell           & 1  \\
\rD_\ell, \ \ell\geqslant 4           & 1    &\ell   & 2\ell          & \text{natural}       & \rA_{\ell-1}              & \ell           & 2  \\ 
\rD_\ell, \ \ell\geqslant 4           & \ell &\ell-1 & 2^{\ell-1}     & \text{half-spinor}   & \rA_{\ell-1}              & \ell           & \ell-2  \\
\rE_\ell, \ \ell=6, 7                  & \ell &2      & 27, 56         & \text{minimal}       & \rA_{\ell-1}              & \ell           & 3       \\ 
\rE_\ell, \ \ell=6, 7                  & \ell &1      & 27, 56         & \text{minimal}       & \rD_{\ell-1}              & 2(\ell-1)      & \ell-1  \\
\rF_4                                & 4    &1     & 26             & \text{minimal}       & \rC_3                     & 6              & 3\end{array}\]
 \caption{List of the cases considered in the proof of \cref{thm:DennisVaserstein}.} \label{table:dv-reps}
\end{table}

Denote by $L_{rs}$ the set consisting of all the elements $g\in \E(\Phi, R, I)$ such that $g \cdot A_{rs} \subseteq A_{rs}$.
It is easy to see that $L_{rs}$ contains $\EP_r(R, I)$ and is normalized by $N_{rs}$, i.\, e. ${L_{rs}}^{N\ssub{rs}}\subseteq L_{rs}$. Indeed, for $g\in N_{rs}$, $h\in L_{rs}$ one has
\begin{equation}\label{rel:NnormL} h^g \cdot A_{rs} = g^{-1} \cdot h \cdot g \cdot A_{rs} \subseteq g^{-1} \cdot h \cdot A_{rs} \cdot g \subseteq {A_{rs}}^g \subseteq A_{rs}.\end{equation}

@@

The next goal is to demonstrate the inclusion $\U(\Sigma_s^+, R) \subseteq N_{rs}$ (which implies $\EP_s(R) \subseteq N_{rs}$).
All the possibilities for $\Phi$, $s$ and $r$ fall into one of the following three cases:
\begin{enumerate}
 \item \textit{Case $m_s(\alpha_{\mathrm{max}})=1$ and $\Phi$ is simply laced.}
  In this case all the roots of $\Sigma^+_s$ have the same $s$-shape and length, hence the assertion follows from \cref{item-trans2}, \eqref{rel:NnormL} and the fact that we already have the inclusion $X_{\alpha_s}(R) \subseteq N_{rs}$.
 \item \textit{Case $\Phi = \rB_\ell$.} Using the argument from the previous case we get that $N_{rs}$ contains the subgroup generated by the root subgroups $X_\alpha(R)$, where $\alpha$ varies over long roots of $\Sigma_s^+$. 
  It remains to prove the inclusion $X_{1, 0}(R) \subset N_{rs}$.
  Specializing identity~\eqref{rel:CCF} we get \begin{equation*} \label{rel:CCF-specBC} x_{1, 0}(ab) = [x_{1, 2}(a), x_{2, 0}(b)] \cdot x_{1, -2}(-a b^2) \end{equation*}
  Since $X_{2, 0}(R)$ and $X_{1, -2}(R)$ are contained in $N_{rs}$, the inclusion $X_{1, 0}(R) \subseteq N_{rs}$ follows.
 \item \textit{Case $\Phi = \rC_\ell, \rF_4$.}
 We settle these remaining cases in \cref{lemma:DVST} where we invoke the stability conditions one more time. 
 In fact, we prove a stronger result, namely that $N_{rs}$ contains $X_{-\alpha\ssub{s}}(R)$ and, consequently, $N_{rs} = \E(\Phi, R)$.
\end{enumerate}

@@

The last inclusion follows from the same transitivity argument as above and the fact that $X_\alpha(I) \subseteq \EP_r(R, I) \subseteq L_{rs}$ for $\alpha \in \Delta_r \cup \Sigma_r$.
For $\Phi=\rB_\ell$ one has to once again apply Chevalley commutator formula~\eqref{rel:CCF}:
\[ X_{0, 1}(I) \subseteq X_{2, 1}(I) \cdot X_{2, 1}(I)^{x_{0, 2}(1)} \cdot X_{-2, 1}(I) \subseteq L_{rs} \cdot L_{rs}^{N_{rs}} \cdot L_{rs} \subseteq L_{rs}. \qedhere \] 
\end{proof}

@@

\begin{lemma}\label{lemma:DVST}
Assume that one of the following holds:
\begin{lemlist}
 \item \label{lemma:DVcaseCl} $\Phi=\rC_\ell$ and $\sr(I) \leq \ell-1$;
 \item \label{lemma:DVcaseF4} $\Phi=\rF_4$ and $\asr(I) \leq 3$.
\end{lemlist}
Then one has $X_{-\alpha_s}(R) \subseteq N_{rs}$.
\end{lemma}
\begin{proof}
Consider the subset $C \subseteq \EP_s(R, I)$ consisting of elements satisfying the following condition:
\begin{itemize}
 \item \textsc{Case $\Phi=\rC_\ell$.} Matrix entries $(g_{i, 2})$, $i=2, \ldots, \ell$ form an $I$-unimodular column of height $\ell-1$.
 \item \textsc{Case $\Phi=\rF_4$.} Matrix entries $(g_{\lambda, \lambda}, \ldots, g_{\lambda, \lambda - \alpha_3 - \alpha_2 - \alpha_1})$, where $\lambda = \varpi_4 - \alpha_4$ (see~\cite[Fig.~26]{PSV98}), 
                                     form an $I$-unimodular column of height $4$. \end{itemize}

@@

By the very definition of $C$, for every $g \in C$ one can choose $y \in \U(\Sigma_s^+ \cap \Delta_r, I)$ such that $(y \cdot g)_{\varpi_s, \varpi_s - \alpha_s} = 0$.
Consequently, for every $g\in C$ one has
\begin{multline*}
 \EP_r(R, I) \cdot \U(\Sigma_s^- \cap \Sigma_r^-, I) \cdot g \subseteq \EP_r(R, I) y^{-1} \cdot \U(\Sigma_r^-\cap \Sigma_s^-, I)^{y^{-1}} \cdot y g \subseteq \\
  \subseteq \EP_r(R, I) \cdot \U(\Sigma_r^-, I) \cdot y g
\end{multline*}
Notice that the matrix entry $(yg)_{\varpi_s, \varpi_s}$ is invertible.
From the choice of $y$ it follows that for every $\xi\in R$ the element $g_1:=(yg)^{x_{-\alpha_s}(\xi)}$
satisfies the assumptions of \cref{lemma:Chevalley-Matsumoto} and therefore can be rewritten as $g_1 = uh$ for some $u \in \U(\Sigma_s^-, I)$, $h \in \EP_1(R, I)$.
Consequently, one has
\begin{multline*} {A_{rs}}^{X_{-\alpha_{s}}(R)} \subseteq \EP_r(R, I)^{X_{-\alpha_{s}}(R)} \cdot \U(\Sigma_r^-, I)^{X_{-\alpha_{s}}(R)} \cdot \U(\Sigma_s^-, I) \cdot \EP_s(R, I) \subseteq \\
 \subseteq \EP_\ell(R, I) \cdot \U(\Phi^-, I) \cdot \EP_s(R, I) \subseteq A_{rs} \end{multline*}
 as claimed.
\end{proof}
